\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-05-14 doc/sw-system.tex
Halfway through intro refs

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{Auger++2011,
Sonnenfeld++2014}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2004,Vegetti+Koopmans2009,Hezaveh++2012}. Any
strong lens can provide magnification of a factor of 10 or more, providing a
deeper, higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{Stark++2009,Newton++2009}. Lensed quasars enable
cosmography via the time delays between the multiple images' lightcurves
\citep[e.g.][]{Tewes++2011,Suyu++2011}, and study of the accretion disk itself
through the microlensing effect \citep[e.g.][]{Morgan++2011}. All of these
science projects would benefit from being able to draw from a larger sample of
lenses.

In the last decade the numbers of detections of these rare cosmic alignments
has increased by an order of magnitude, thanks to wide field surveys such as
CLASS \citep[e.g.]{Browne++2003}, SDSS \citep[e.g.][]
{Bolton++2006,Auger++2010,Treu++2011,Oguri++2010}, CFHTLS
\citep[e.g.][]{More++2012,Gavazzi++2014}, Herschel \citep[][]{Herschel} and
SPT \citep[e.g.][]{Vieira++2012}, among others.  As the number of known lenses
has increased, new types have been discovered, leading to entirely new
investigations. Compound lenses \citep{Gavazzi++2007,Collett+Auger2014} and
lensed supernovae \citep{Quimby++2014} are good examples of this.

@@

Because they are rare, strong lenses are expensive to find. The most efficient
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{Bolton++2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{Negrello++2010}. Such searches have to be efficient,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields in the tens to hundreds. An alternative approach
is to search images of sufficiently high resolution and color contrast, and
confirm the systems as gravitational lenses by modeling the survey data
themselves \citep[][]{Marshall++2009}. Several square degrees of HST images
have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{Moustakas++2006,Faure++2009,Jackson++2010,Pawase++2012}.
Similarly, searches of over a hundred square degrees of CFHT Legacy Survey
ground-based imaging, also with sub-arcsecond image quality, have revealed a
smaller number of wider image separation group-scale systems
\citep[e.g.][]{Cabanac++2007,More++2010}. Detecting galaxy-scale lenses from
the ground is hard, but feasible albeit lower efficiency and requiring HST or
spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{Gavazzi++2014}.

@@

How can we scale these lens searches up to imaging surveys covering a hundred
times the sky area, such as the almost-all sky surveys planned with LSST and
Euclid, while reducing our dependence on expensive follow-up confirmation
observations? There are two approaches to detecting lenses in imaging surveys.
The first one is robotic: automated analysis of object catalogs and/or the
survey images. The candidate samples produced by these methods have, to date,
not been of high purity  \citep[see
e.g.][]{Marshall++2009,More++2010,Gavazzi++2014}, with visual inspection by
teams of humans still required to narrow down the robotically-generated
samples. In this approach, the image data may or may not be explicitly
modelled by the robots as if it contained a gravitational lens, but the visual
inspection can be thought of as a ``mental modeling'' step. Systems classified
by an inspector to be good lens candidates are deemed as such because the
features in the image can be explained by a model of what gravitational lenses
do contained in the inspector's brain. The second approach simply cuts out the
robot middleman: \citet{Faure++2009,Jackson++2010} and \citet{Pawase++2012}
all performed entirely visual searches for lenses in HST imaging.

>>>>

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{AugerEtal2010,
SonnenfeldEtal2012,SonnenfeldEtal2013}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2002,VegettiEtal2010,HezavehEtal2013}. Any
strong lens can provide magnification of a factor of 10 or more, providing a
deeper, higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{StarkEtal2008,NewtonEtal2011}. Lensed quasars enable
cosmography via the time delays between the multiple images' lightcurves
\citep[e.g.][]{TewesEtal2013,SuyuEtal2013}, and study of the accretion disk itself
through the microlensing effect \citep[e.g.][]{PoindexterEtal2008}. All of these
science projects would benefit from being able to draw from a larger sample of
lenses.

In the last decade the numbers of detections of these rare cosmic alignments
has increased by an order of magnitude, thanks to wide field surveys such as
CLASS \citep[e.g.]{BrowneEtal2003}, SDSS \citep[e.g.][]
{BoltonEtal2006,AugerEtal2010b,TreuEtal2011,InadaEtal2012}, CFHTLS
\citep[e.g.][]{MoreEtal2012,GavazziEtal2014}, Herschel \citep[][]{Herschel} and
SPT \citep[e.g.][]{VieiraEtal2013}, among others.  As the number of known lenses
has increased, new types have been discovered, leading to entirely new
investigations. Compound lenses \citep{GavazziEtal2008,Collett+Auger2012} and
lensed supernovae \citep{QuimbyEtal2014} are good examples of this.

@@

Because they are rare, strong lenses are expensive to find. The most efficient
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to be efficient,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields in the tens to hundreds. An alternative approach
is to search images of sufficiently high resolution and color contrast, and
confirm the systems as gravitational lenses by modeling the survey data
themselves \citep[][]{MarshallEtal2009}. Several square degrees of HST images
have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2006,FaureEtal2009,JacksonEtal2010,PawaseEtal2012}.
Similarly, searches of over a hundred square degrees of CFHT Legacy Survey
ground-based imaging, also with sub-arcsecond image quality, have revealed a
smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2010}. Detecting galaxy-scale lenses from
the ground is hard, but feasible albeit lower efficiency and requiring HST or
spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

@@

How can we scale these lens searches up to imaging surveys covering a hundred
times the sky area, such as the almost-all sky surveys planned with LSST and
Euclid, while reducing our dependence on expensive follow-up confirmation
observations? There are two approaches to detecting lenses in imaging surveys.
The first one is robotic: automated analysis of object catalogs and/or the
survey images. The candidate samples produced by these methods have, to date,
not been of high purity  \citep[see
e.g.][]{MarshallEtal2009,MoreEtal2010,GavazziEtal2014}, with visual inspection by
teams of humans still required to narrow down the robotically-generated
samples. In this approach, the image data may or may not be explicitly
modelled by the robots as if it contained a gravitational lens, but the visual
inspection can be thought of as a ``mental modeling'' step. Systems classified
by an inspector to be good lens candidates are deemed as such because the
features in the image can be explained by a model of what gravitational lenses
do contained in the inspector's brain. The second approach simply cuts out the
robot middleman: \citet{FaureEtal2009,JacksonEtal2010} and \citet{PawaseEtal2012}
all performed entirely visual searches for lenses in HST imaging.
